Calendar-based graphics for visualizing people's daily schedules

### 2017-08-29 main.Rmd
emphasised on fitting into the tidyverse framework intead of extension of ggplot2

keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, R package, grammar of graphics]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:

>>>>

keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, and fits into the \pkg{tidyverse} [@R-tidyverse] and grammar of graphics [@wickham2009ggplot2] framework for exploring data. The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. The sensor data, like many temporal datasets on human behaviors, lends itself to a number of exploratory data visualization challenges:

### 2017-08-29 main.tex
emphasised on fitting into the tidyverse framework intead of extension of ggplot2

\Keywords{data visualization, statistical graphics, time series, grammar of graphics, \proglang{R}}
\Plainkeywords{data visualization, statistical graphics, time series, R package, grammar of graphics}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The purpose of
the calendar-based visualization is to provide insights into people's
daily schedules relative to events such as work days, weekends,
holidays, and special events. This work was originally motivated by
studying foot traffic in the city of Melbourne, Australia \citep{ped}.
There have been 43 sensors installed that count pedestrians every hour
across the downtown area (Figure \ref{fig:ped-map}). The dataset can
shed light on people's daily rhythms, and assist the city administration
and local businesses with event planning and operational management. A
routine examination of the data would involve constructing conventional
time series plots to catch a glimpse of temporal patterns. The faceted
plots in Figure \ref{fig:time-series-plot}, give an overall picture of
the foot traffic at 3 different sensors over 2016. Further faceting by
day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
of the daily and sub-daily pedestrian patterns. The sensor data, like
many temporal datasets on human behaviors, lends itself to a number of
exploratory data visualization challenges:

>>>>

\Keywords{data visualization, statistical graphics, time series, grammar of graphics, \proglang{R}}
\Plainkeywords{data visualization, statistical graphics, time series, grammar of graphics, R}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, and fits into the
\pkg{tidyverse} \citep{R-tidyverse} and grammar of graphics
\citep{wickham2009ggplot2} framework for exploring data. The purpose of
the calendar-based visualization is to provide insights into people's
daily schedules relative to events such as work days, weekends,
holidays, and special events. This work was originally motivated by
studying foot traffic in the city of Melbourne, Australia \citep{ped}.
There have been 43 sensors installed that count pedestrians every hour
across the downtown area (Figure \ref{fig:ped-map}). The dataset can
shed light on people's daily rhythms, and assist the city administration
and local businesses with event planning and operational management. A
routine examination of the data would involve constructing conventional
time series plots to catch a glimpse of temporal patterns. The faceted
plots in Figure \ref{fig:time-series-plot}, give an overall picture of
the foot traffic at 3 different sensors over 2016. Further faceting by
day of the week (Figure \ref{fig:facet-time}) provides a better glimpse
of the daily and sub-daily pedestrian patterns. The sensor data, like
many temporal datasets on human behaviors, lends itself to a number of
exploratory data visualization challenges:
